---
title: "Report"
author: "Bryana Gutierrez, Lily Li, Erica Wong"
date: "December 8, 2015"
output: pdf_document
---

```{r, echo = FALSE}
library (knitr)
read_chunk("../code/cleaning_data.R")
read_chunk("../code/part_1.R")
read_chunk("../code/part_2.R")
read_chunk("../code/part_3.R")
```
## Cleaning Data
This script starts off with two raw data files and we first worked on cleaning
each individually by extracting the data we wanted and then concatenated both 
data sets based on the date of the accidents. 

First we download the raw data onto R studio: 
```{r Loading Data, cache = TRUE}
```

We will use a numbering convention so that at each modification the index 
increases by one. This takes only the airplane data from the data frame as 
opposed to all aircrafts.
```{r Cleaning1, cache = TRUE}
```

Next, we merge these two date tables so that air_data provides information
from 1920-1981 and av_data information from 1982-2015. So now we extract 
only the years we want from each data frame. 
```{r Cleaning2, cache = TRUE}
```

In av_data the columns 'Location' and 'Country' should be combined as well
as 'Make' and 'Model'
```{r Cleaning3, cache = TRUE}
```

Now, we work on reorganizing the data so that it is in the format that we 
want. In air_data the order of data is currently in ascending order, but 
to match av_data we want it to be in descending order. We also extract the
columns that we want and rename then so that the two data frames have the same
column names. We do this, so that once we extract missing values we can 
concatenate the two data frames. Finally we make the type and location names 
uniform by changing all characters to lower case. 
```{r Aesthetics, cache = TRUE}
```

This is our clean data
```{r Peek, cache = TRUE}
```

Exporting the `data` file into the data folder. 
```{r Exporting, cache = TRUE}
```

## Part 1: Monthly Data
In this part of the project, we were looking to see which of the months had 
the most plane crashes. In order to do this, we will need to first load the 
necessary code and packages.
```{r Necessary Packages and Functions Part 1, cache = TRUE}
```

Our cleaned data set does not have a column that just contains the months when
each accident happens, so we will had to make our own by doing the following.
```{r Adding Month Column, cache = TRUE}
```

With this column, we are now ready to graph to see our findings.
```{r Graphing Monthly Data, cache = TRUE}
```

Our original belief was that the number of accidents every month would be 
relatively uniform, but instead, we found that there are more accidents that 
occur during the middle of the year, with the most occurring in July. We found 
this idea to be interesting and surprising. The number of accidents could have 
increased due to there being more traveling being done during the summer due to
kids being on summer vacation. However, during the summer seasons, there are 
also a lot of tropical storms , thunderstorms, and changes in the weather that
could affect how safe it is to travel and increases the chance of getting into 
a plane accident. According to a New York Times article, the number of airplane 
accidents could be greater in the summer months because that is typically when
people fly their private airplanes. A smaller, private aircraft is more likely
to get into an accident than a commercial plane. The article states that, 
“Statistics from the N.T.S.B. show that general aviation aircraft average 
nearly seven accidents per 100,000 flight hours, compared with an average of
0.16 accidents per 100,000 hours for commercial airlines” (Fowler). 

Finally, we will export our graphics.
```{r Exporting Graphics Part 1, cache = TRUE}
```

## Part 2: Data by Decades

```{r Necessary Packages and Functions, cache = TRUE}
```

We wanted to look at the number of accidents per decade to see if any 
particular decades differed from the rest. However, the data is not organized 
into decades. First, we used regular expressions to pull the year out of the 
date column and created a new column named "year". 

```{r New Numeric Year Column, cache = TRUE}
```

For graphing, we wanted the axis to display decades, for example "1980s", and
the corresponding aggregated data for that decade. We created a loop that 
categorizes the year and would group all years in a certain decade into that
decade. 

```{r Creating Decades Labels for each Data Point, cache = TRUE}
```

Next, we looked at the frequency of accidents for each decade by calling upon
the data through table(). 

```{r Amount of Accidents in each Decade}
```

We started off by graphing a bar chart of the number of accidents per decade 
but found that the data was skewed due to the significantly larger size of the 
dataset for years 1980s and onwards. We then normalized the data by taking 
proportion of each decade’s accidents to the total amount of accidents for the
data set it came from. For example, 1920s proportion was: 
(# of accidents in 1920s) / (total # of accidents of dataset1)

```{r Normalizing Data, cache = TRUE}
```

There seems to be a gradual increase in accidents till 1980s where it starts to 
drop off starting 1990s. Data from the 2000s show a decrease, which agrees with
the FAA report. 

```{r Graphing Proportions of Accidents of each Decade, cache = TRUE}
```

According to the FAA:
“Between 2001 and 2007, aviation witnessed one of its safest periods for 
scheduled air carriers. Not counting the terrorist activities of September 11, 
2001, there were only three fatal accidents in 2001; none in 2002; two in 2003; 
one in 2004; three in 2005; two in 2006; and none in 2007. Fatal accidents 
became rare events with only .01 accidents per 100,000 flight hours or .018 
accidents per 100,000 departures.”

Another explanation could be the improvements of air traffic control since 
after the 1980s (start of the digital age). 

According to Wikipedia:
As computers became more sophisticated in the 2000s, they began to take over 
routine aspects of the air traffic controller's task. Up until then all air 
traffic in nearby airspace was tracked and displayed, with the air traffic 
controller responsible for monitoring its position and assessing any need for 
action. Modern computerized systems are capable of monitoring the flight paths
of many more aircraft at a given time, allowing the controller to manage more
aircraft and to focus on the decision-making and follow-up processes.

```{r Exporting Graphics, cache = TRUE}
```

## Part 3: Type analysis 
Here we are looking at how the most common types of airplanes to crash changed 
over different decades. We are parse out the data by decade so that we can plot
by type. Specifically, we extract just the top three values.
```{r Preliminary, cache = TRUE}
```

```{r Top3 Array, cache = TRUE}
```

Now we want to create a data frame that combines the data from each decade in 
a way that we can plot it efficiently. 
```{r Reorganizing, cache = TRUE}
```

These are the two sets of decades we want to focus on. Because of the 
differences among our data sets, we want to look at:
  - from 1940-1980 
```{r Parsing: 1940-1980, cache = TRUE}
```

From our graph, we can see that from 1940-1979, the Douglas DC-3, was the plane
that got into the most accidents. However, this could be due to the fact that 
the Douglas DC-3 was the most produced airliner. So, because there were more of
them in production, the chances that they got into accidents were higher for 
that reason.

  - from 1980-2010
```{r Parsing: 1980-2010, cache = TRUE}
```

From our graph here, we can see that from 1980-2010 the plane that got into the 
most accidents was Cessina 152. In fact, all of the top 3 planes that got into
accidents are Cessina. Cessina 152 is a private plane and according to an 
article from The Economist, commercial air travel has been a lot safer since
the 1980s, which would explain why from 1980 forward, private planes, such as 
Cessina, have been getting into more accidents. There are also many other
reasons as to why Cessina and other private planes have been getting into more 
accidents. For example, not a lot of private plane owners have a lot of 
experience, private planes are flown by one person and most accidents have come
from them losing control or not seeing something and flying into it due to no 
one looking out for them, and having smaller fuel capacities so there is less 
room for error. Finally, we might have seen more Cessinas involved in 
accidents because there have been more Cessina produced in recent years than 
any other kind of plane. 

Finally, We export the graphics. 
```{r Exporting the Graphics, echo = FALSE, cache = TRUE}
```
